Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-07-26 paper/ms.tex
update ms

An interesting point of comparison is the rapidly rotating magnetic
B-star, $\sigma$~Ori~E, which shows dips that are deeper in the blue
than in the red \citep{1977ApJ...216L..31H}.  Photometric and
spectroscopic observations of this star have been understood in terms
of absorbing circumstellar material, analogous to the geometries we
are discussing for much cooler M-dwarfs \citep{2005ApJ...630L..81T}.
This material however is unlikely to be dust, due to the relevant
sublimation timescales.  The opacity is instead thought to be sourced
from bound-free absorption from neutral hydrogen
\citep{2005ApJ...630L..81T}.  Separate and smaller-amplitude emission
in that system may also come from electrons scattering photospheric
light toward the observer when the clouds are not in transit
\citep{2022MNRAS.511.4815B}.

@@

\subsection{Strengthening the magnetic B-Star connection}
Previous work by \citet{2017AJ....153..152S} noted an intriguing
connection between the CQVs and rapidly rotating magnetic B-stars such
as $\sigma$~Ori~E, which are known to have circumstellar gas clouds
trapped in corotation \citep{2005ApJ...630L..81T}.  Spectropolarimetry
has indicated that, similar to $\sigma$~Ori~E, most magnetic B-stars
have relatively simple field geometries that are predominantly
dipolar, though generally not aligned with the stellar spin axis
\citep{2009ARA&A..47..333D}.  Two exceptions to this rule are HD~37776
and HD~64740, which are B-stars that were previously known from
spectropolarimetry in the 2000s-2010s to have more complicated field
geometries - dominated by high order multipoles
\citep{2011ApJ...726...24K}.  It turns out, recent TESS light curves
of these two B-stars look surprisingly similar to the CQV light curves
\citep{2020pase.conf...46M}.  
Figure~\ref{fig:bstar} shows the phased TESS light curves for
the two outlying magnetic B-stars,
with by-eye ``best matching'' CQVs shown for comparison.
The number of dips per cycle, the shapes of the dips, and the dip
depths relative to the sinusoidal envelope, all seem to be quite
similar.
If this connection is taken seriously, then the implication would be
that the highly-structured M-dwarf light curves imply multipolar magnetic fields --
since that is the defining characteristic of these two B-stars,
relative to other known magnetic B stars like $\sigma$~Ori~E.

>>>>

An instructive point of comparison is the rapidly rotating magnetic
B-star, $\sigma$~Ori~E, which shows dips that are deeper in the blue
than in the red \citep{1977ApJ...216L..31H}.  Photometric and
spectroscopic observations of this star have been understood in terms
of a warped torus of circumstellar material, analogous to the
geometries we are discussing for much cooler M-dwarfs
\citep{1978ApJ...224L...5L,2005ApJ...630L..81T}.  This material
however is unlikely to be dust, due to the relevant sublimation
timescales.  The opacity is instead thought to be sourced from
bound-free absorption from neutral hydrogen
\citep{2005ApJ...630L..81T}, although to our best knowledge direct
evidence for this conclusion has yet to be acquired.  Separate and
smaller-amplitude emission in $\sigma$~Ori~E may also come from
electrons scattering photospheric light toward the observer when the
clouds are not in transit \citep{2022MNRAS.511.4815B}.

@@

\subsection{Strengthening the magnetic B-star connection}

Previous work by \citet{2017AJ....153..152S} noted the possible
connection between the CQVs and rapidly rotating magnetic B-stars such
as $\sigma$~Ori~E, which can have circumstellar gas clouds trapped in
corotation \citep{2005ApJ...630L..81T}.  The $\sigma$~Ori class is
distinct from Be-star decretion disks, which do not have any obvious
connection to the stellar magnetic field \citep{2013A&ARv..21...69R}.

One possible argument against the connection between CQVs and the
$\sigma$~Ori~E analogs is that the light curve of $\sigma$~Ori~E is
much simpler than those in Figure~\ref{fig:cqvs}, with only two broad
local minima, and one ``hump'' \citep{2005ApJ...630L..81T}.  Within
the model proposed by Townsend et al., this is tied to the star's
magnetic field, which to first-order is reasonably described as a
dipole, similar to most other magnetic B-stars
\citep{2009ARA&A..47..333D}.  This approximation fits the qualitative
behavior of not only the broadband light curves, but also the
line-profile variations seen in hydrogen, helium, and carbon
\citep{2012MNRAS.419..959O}.

Two exceptions to the rule of ``simple field topologies for magnetic
B-stars'' are HD~37776 and HD~64740, which are B-stars that were
previously known from spectropolarimetry in the 2000s-2010s to have
field geometries dominated by high order multipoles
\citep{2011ApJ...726...24K}.  In a review of the literature, we
learned that recent TESS light curves of these two B-stars appear
surprisingly similar to the CQV light curves
\citep{2020pase.conf...46M}.  Figure~\ref{fig:bstar} shows the phased
TESS light curves for the two outlying magnetic B-stars, with by-eye
``best matching'' CQVs shown for comparison.  The number of dips per
cycle, the shapes of the dips, and the dip depths relative to the
sinusoidal envelope, all seem to be quite similar.  If this connection
is taken seriously, the implication would be that the
highly-structured M-dwarf light curves imply multipolar magnetic
fields -- since it is the complex fields which are the defining
characteristic of these two B-stars, relative to other known magnetic
B stars like $\sigma$~Ori~E.  
This is consistent with the non-axisymmetric topologies that have been
reported in a small number of M-dwarfs for which Zeeman Doppler
Imaging is technically feasible 
\citep[see the recent review by][]{2021A&ARv..29....1K}.

The physical similarity between the B-stars and the M-dwarfs is
presumably the existence of a ``centrifugal magnetosphere'': in other
words, they both satisfy the condition $R_{\rm A} > R_{\rm cr}$, for
$R_{\rm A}$ the Alfv\'en radius.
% TODO : qualitiatively explain, and cite Petit+2013
In the converse case of a ``dynamical'' magnetosphere...
{\bf todo: summarize Petit2013}
%FIXME TODO
%FIXME TODO
%FIXME TODO
{\bf
LGB todo: make the connection to Jupiter-Io magnetosphere, and that
plasma torus too
}
